---
title: "OpenField"
output:   
  html_document:
    code_folding: hide 
    number_sections: true
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Libraries
```{r}
library(readxl)
library(tidyverse)
library(psych)
library(DescTools)
library(car)
library(stats)
library(broom)
library(dunn.test)
library(stats)
library(ggpubr)
library(rstatix)
```


# Loadind data
```{r}

data <- read_excel("Rotarod.xlsx")
head(data)

```

```{r}
data <- data %>% mutate(Idmice=as.factor(Idmice),
                        Line=as.factor(Line),
                        Genotype=as.factor(Genotype),
                        Group=as.factor(Group),
                        Gender=as.factor(Gender))
```


```{r}
# Function to clean column names
clean_column_names <- function(x) {
  gsub(" ", "_", gsub("-", "_", gsub("___","_",x)))
}

# Apply the clean_column_names function to column names of the data frame
colnames(data) <- clean_column_names(colnames(data))
colnames(data) <- clean_column_names(colnames(data))#Run twice to complete conversion. It can be optimized.

```

```{r}
colnames(data)
# Convert "Group" to factor with specified order
data$Group <- factor(data$Group, levels = c("NTC", "WT", "L115R", "L116"))
str(data)
```



# Functions

```{r}
# Function to convert p-values to star scale
convert_to_star <- function(p) {
  if (p > 0.05) {
    return("ns")
  } else if (p <= 0.05 & p > 0.01) {
    return("*")
  } else if (p <= 0.01 & p > 0.001) {
    return("**")
  } else {
    return("***")
  }
}

# Function to clean up variable names for file paths
clean_variable_name <- function(var_name) {
  cleaned_name <- gsub("[[:punct:]]", "_", var_name)
  return(cleaned_name)
}


```


# ANOVA

## Grouping by Gender 
### Homocedasticity by Gender

```{r}
aovdata <- data[, c(3,6, 8:17)]
variables <- colnames(data[,8:17])
# Assuming your dataframe is called 'aovdata', and the 'Gender' column contains the gender information
# and all the numeric variables you mentioned are in the dataframe.

# Filter numeric variables (excluding 'Group' and 'Gender' columns)
numeric_vars <- names(aovdata)[!(names(aovdata) %in% c("Group", "Gender"))]

# Gather the numeric variables into a tidy format
aovdata_tidy <- aovdata %>%
  select(Gender, all_of(numeric_vars)) %>%
  pivot_longer(cols = -Gender, names_to = "Variable", values_to = "Value")

# Convert the 'Gender' column to a factor (if it's not already)
aovdata_tidy$Gender <- as.factor(aovdata_tidy$Gender)

# Perform Levene test by gender for each numeric variable
levene_results <- aovdata_tidy %>%
  group_by(Variable) %>%
  summarise(Levene_p_value = list(leveneTest(Value, group = Gender)$`Pr(>F)`), .groups = 'drop') %>%
  unnest(Levene_p_value)
# Filter out rows with NA in the Levene_p_value column
levene_results_gender <- levene_results %>%
  filter(!is.na(Levene_p_value))%>% 
  arrange(Levene_p_value)


levene_results_gender$StarScale <- sapply(levene_results_gender$Levene_p_value, convert_to_star)
# View the results
print(levene_results_gender)
```
The results of Levene Test by Gender said us that there are not differences in homogeneity of samples. We can apply the ANOVA test.  Let's see Normality

### Normality by Gender 

```{r}
# Perform Shapiro-Wilk test for each numeric variable grouped by "Group" and "Gender"
aovdata <- data[, c(3,6, 8:17)]
# Group-wise Shapiro-Wilk test
shapiro_test_gender <- aovdata %>%
  group_by(Gender) %>%
  summarize(across(where(function(x) length(unique(x)) > 1 && is.numeric(x)), ~ shapiro.test(.)$p.value, .names = "{.col}")) %>%
  ungroup()
shapiro_test_gender<-as.data.frame(t(shapiro_test_gender))

# Rename columns V1 and V2 to Females and Males
colnames(shapiro_test_gender)[colnames(shapiro_test_gender) == "V1"] <- "Females"
colnames(shapiro_test_gender)[colnames(shapiro_test_gender) == "V2"] <- "Males"

# Format the values in scientific notation to regular numeric format
shapiro_test_gender$Females <- format(as.numeric(shapiro_test_gender$Females), scientific = FALSE)
shapiro_test_gender$Males <- format(as.numeric(shapiro_test_gender$Males), scientific = FALSE)
# Add a new column for the test result
shapiro_test_gender$Significance_Females <- ifelse(shapiro_test_gender$Females < 0.05, "Non-Normal", "Normal")
shapiro_test_gender$Significance_Males <- ifelse(shapiro_test_gender$Males < 0.05, "Non-Normal", "Normal")
# Remove the "Gender" row
shapiro_test_gender <- shapiro_test_gender[!(row.names(shapiro_test_gender) %in% c("Gender")),]
# Arrange the values by Females in ascending order
shapiro_test_gender <- shapiro_test_gender[order(shapiro_test_gender$Females), ]
# Add row names as a new column to the data frame
shapiro_test_gender <- shapiro_test_gender %>% rownames_to_column(var = "Variable")

shapiro_test_gender$StarScaleMale <- sapply(shapiro_test_gender$Males, convert_to_star)
shapiro_test_gender$StarScaleFemale <- sapply(shapiro_test_gender$Females, convert_to_star)
# View the result
print(shapiro_test_gender)

#Group variables if pass or not normality

# Get variables that are not significant for both Males and Females
significant_males <- shapiro_test_gender %>%
  filter(Significance_Males == "Non-Normal") %>% 
  pull(Variable)
# Get variables that are not significant for both Males and Females
not_significant_males <- shapiro_test_gender %>%
  filter(Significance_Males == "Normal") %>% 
  pull(Variable)
# Get variables that are not significant for both Males and Females
significant_females <- shapiro_test_gender %>%
  filter(Significance_Females == "Non-Normal") %>% 
  pull(Variable)
# Get variables that are not significant for both Males and Females
not_significant_females <- shapiro_test_gender %>%
  filter(Significance_Females == "Normal") %>% 
  pull(Variable)
```



## Males ANOVA

```{r}
# Load required libraries (make sure you have installed 'broom' and 'dplyr' packages)
library(dplyr)
library(broom)

```


```{r}
# Assuming your original data frame is named 'data', select relevant columns
aovdata <- data[, c(3,6, 8:17)]

# Convert "Group" and "Gender" to factors
aovdata$Group <- as.factor(aovdata$Group)
aovdata$Gender <- as.factor(aovdata$Gender)

# Initialize an empty list to store ANOVA or Kruskal-Wallis results
male_aovlist <- list()

male_data <- aovdata %>% filter(Gender == "Male")

for (col in names(male_data)[3:ncol(male_data)]) {
      variable <- col
    formula <- as.formula(paste0("`", col, "`"," ~ Group"))
  
  # Check for normality assumption using Shapiro-Wilk test
  if (shapiro.test(male_data[[col]])$p.value >= 0.05) {
    #Include Levene Test here
    res.aov <- aov(formula, data = male_data)
    data_input <- data.frame("Group"=male_data$Group, "Gender"=male_data$Gender,male_data[[col]])
    names(data_input)[3]<- variable
    data_input$Group <- as.factor(data_input$Group)
    posthoc <- data_input %>% rstatix::tukey_hsd(formula) # Tukey's post-hoc test for ANOVA
    posthoc<- posthoc %>% add_x_position(dodge = 0.8)
    posthoc<- posthoc %>% add_y_position()
    test_type <- "ANOVA"
    Group1 <- posthoc$group1
    Group2 <- posthoc$group2
    posthoc$xmin<- posthoc$xmin+0.2
    posthoc$xmax<- posthoc$xmax+0.2
    
    #pair_compared <- rownames(posthoc$`Group`)
  } else {
    res.aov <- kruskal.test(formula, data = male_data)
    data_input <- data.frame("Group"=male_data$Group, "Gender"=male_data$Gender,male_data[[col]])
    names(data_input)[3]<- variable
    data_input$Group <- as.factor(data_input$Group)
    posthoc <- data_input %>% dunn_test(formula)
    test_type <- "Kruskal-Wallis"
    Group1 <- posthoc$group1
    Group2 <- posthoc$group2
    posthoc<- posthoc %>% add_x_position(dodge = 0.8)
    posthoc<- posthoc %>% add_y_position()
    posthoc$xmin<- posthoc$xmin+0.2
    posthoc$xmax<- posthoc$xmax+0.2
    #pair_compared <- posthoc$comparison
  }
  male_aovlist[[col]] <- list(aov_results = broom::tidy(res.aov, conf.int = TRUE), posthoc_results = posthoc, Test_Type = test_type, Variable = variable)
}
```


```{r}
# Create empty data frame to store the results
male_aov_results_df <- data.frame()

# Loop through the list of test results
for (i in seq_along(male_aovlist)) {
  test_result <- male_aovlist[[i]]
   # Extract variable name
  variable <- test_result$Variable
  # Extract test type
  test_type <- test_result$Test_Type
  # Extract comparison and p-values
  if (grepl("ANOVA", test_type, ignore.case = TRUE)) {
    comparisons <- data.frame("Comparison"=paste0(test_result$posthoc_results$group1,"-",test_result$posthoc_results$group2),
                              "p_value_adj"=test_result$posthoc_results$p.adj,
                              "p.adj.signif"=test_result$posthoc_results$p.adj.signif,
                              "y.position"=test_result$posthoc_results$y.position,
                              "xmin"= test_result$posthoc_results$xmin,
                              "xmax"= test_result$posthoc_results$xmax)
    p_values <- test_result$aov_results$p.value[1]
    
    
  } else if (grepl("Kruskal-Wallis", test_type, ignore.case = TRUE)) {
    comparisons <- data.frame("Comparison"=paste0(test_result$posthoc_results$group1,"-",test_result$posthoc_results$group2),
                              "p_value_adj"=test_result$posthoc_results$p.adj,
                              "p.adj.signif"=test_result$posthoc_results$p.adj.signif,
                              "y.position"=test_result$posthoc_results$y.position,
                              "xmin"= test_result$posthoc_results$xmin,
                              "xmax"= test_result$posthoc_results$xmax)
    p_values <- test_result$aov_results$p.value
  }
  
  # Combine the results into a data frame
  male_aov_results_df <- rbind(male_aov_results_df, data.frame(
    Variable = variable,
    Test_Type = test_type,
    Comparison = comparisons,
    P_Value = p_values
  ))
}
#Add StarScale
#male_aov_results_df$StarScale <- sapply(male_aov_results_df$Comparison.p_value_adj, convert_to_star)
#Add gender column
male_aov_results_df$Gender <- "Male"


# Usamos strsplit para dividir la columna por el carácter "-" o " - "
split_data <- strsplit(male_aov_results_df$Comparison.Comparison, split = "-")


# Extraemos los elementos de la lista resultante y los asignamos a las nuevas columnas
male_aov_results_df$group1 <- sapply(split_data, function(x) x[[1]])
male_aov_results_df$group2 <- sapply(split_data, function(x) x[[2]])
#change names of columns
colnames(male_aov_results_df) <- gsub("Comparison\\.", "", colnames(male_aov_results_df))

male_aov_results_df<- male_aov_results_df %>% mutate_if(is.character, as.factor)
# Print the dataframe
print(male_aov_results_df)

# Filter variables with significant differences between groups based on the ANOVA results (alpha level = 0.05)
male_significant_vars <- subset(male_aov_results_df, p_value_adj < 0.05)
new_row_names <- gsub("\\.1$", "", rownames(male_significant_vars))
rownames(male_significant_vars) <- new_row_names

print(male_significant_vars)

```




## Females ANOVA

```{r}
# Assuming your original data frame is named 'data', select relevant columns
aovdata <- data[, c(3,6, 8:17)]

# Convert "Group" and "Gender" to factors
aovdata$Group <- as.factor(aovdata$Group)
aovdata$Gender <- as.factor(aovdata$Gender)

# Initialize an empty list to store ANOVA or Kruskal-Wallis results
female_aovlist <- list()

female_data <- aovdata %>% filter(Gender == "Female")

for (col in names(female_data)[3:ncol(female_data)]) {
      variable <- col
    formula <- as.formula(paste0("`", col, "`"," ~ Group"))
  
  # Check for normality assumption using Shapiro-Wilk test
  if (shapiro.test(female_data[[col]])$p.value >= 0.05) {
    #Include Levene Test here
    res.aov <- aov(formula, data = female_data)
    data_input <- data.frame("Group"=female_data$Group, "Gender"=female_data$Gender,female_data[[col]])
    names(data_input)[3]<- variable
    data_input$Group <- as.factor(data_input$Group)
    posthoc <- data_input %>% rstatix::tukey_hsd(formula) # Tukey's post-hoc test for ANOVA
    posthoc<- posthoc %>% add_xy_position(dodge = 0.8)
    test_type <- "ANOVA"
    Group1 <- posthoc$group1
    Group2 <- posthoc$group2
    posthoc$xmin<- posthoc$xmin-0.2
    posthoc$xmax<- posthoc$xmax-0.2
    #pair_compared <- rownames(posthoc$`Group`)
  } else {
    res.aov <- kruskal.test(formula, data = female_data)
    data_input <- data.frame("Group"=female_data$Group, "Gender"=female_data$Gender,female_data[[col]])
    names(data_input)[3]<- variable
    data_input$Group <- as.factor(data_input$Group)
    posthoc <- data_input %>% dunn_test(formula)
    test_type <- "Kruskal-Wallis"
    Group1 <- posthoc$group1
    Group2 <- posthoc$group2
    posthoc<- posthoc %>% add_xy_position(dodge = 0.8)
    posthoc$xmin<- posthoc$xmin-0.2
    posthoc$xmax<- posthoc$xmax-0.2
    #pair_compared <- posthoc$comparison
  }
  female_aovlist[[col]] <- list(aov_results = broom::tidy(res.aov, conf.int = TRUE), posthoc_results = posthoc, Test_Type = test_type, Variable = variable)
}
```


```{r}
# Create empty data frame to store the results
female_aov_results_df <- data.frame()

# Loop through the list of test results
for (i in seq_along(female_aovlist)) {
  test_result <- female_aovlist[[i]]
   # Extract variable name
  variable <- test_result$Variable
  # Extract test type
  test_type <- test_result$Test_Type
  # Extract comparison and p-values
  if (grepl("ANOVA", test_type, ignore.case = TRUE)) {
    comparisons <- data.frame("Comparison"=paste0(test_result$posthoc_results$group1,"-",test_result$posthoc_results$group2),
                              "p_value_adj"=test_result$posthoc_results$p.adj,
                              "p.adj.signif"=test_result$posthoc_results$p.adj.signif,
                              "y.position"=test_result$posthoc_results$y.position,
                              "xmin"= test_result$posthoc_results$xmin,
                              "xmax"= test_result$posthoc_results$xmax)
    p_values <- test_result$aov_results$p.value[1]
    
    
  } else if (grepl("Kruskal-Wallis", test_type, ignore.case = TRUE)) {
    comparisons <- data.frame("Comparison"=paste0(test_result$posthoc_results$group1,"-",test_result$posthoc_results$group2),
                              "p_value_adj"=test_result$posthoc_results$p.adj,
                              "p.adj.signif"=test_result$posthoc_results$p.adj.signif,
                              "y.position"=test_result$posthoc_results$y.position,
                              "xmin"= test_result$posthoc_results$xmin,
                              "xmax"= test_result$posthoc_results$xmax)
    p_values <- test_result$aov_results$p.value
  }
  
  # Combine the results into a data frame
  female_aov_results_df <- rbind(female_aov_results_df, data.frame(
    Variable = variable,
    Test_Type = test_type,
    Comparison = comparisons,
    P_Value = p_values
  ))
}
#Add StarScale
#female_aov_results_df$StarScale <- sapply(female_aov_results_df$Comparison.p_value_adj, convert_to_star)
#Add gender column
female_aov_results_df$Gender <- "Female"


# Usamos strsplit para dividir la columna por el carácter "-" o " - "
split_data <- strsplit(female_aov_results_df$Comparison.Comparison, split = "-")


# Extraemos los elementos de la lista resultante y los asignamos a las nuevas columnas
female_aov_results_df$group1 <- sapply(split_data, function(x) x[[1]])
female_aov_results_df$group2 <- sapply(split_data, function(x) x[[2]])
#change names of columns
colnames(female_aov_results_df) <- gsub("Comparison\\.", "", colnames(female_aov_results_df))

female_aov_results_df<- female_aov_results_df %>% mutate_if(is.character, as.factor)
# Print the dataframe
print(female_aov_results_df)

# Filter variables with significant differences between groups based on the ANOVA results (alpha level = 0.05)
female_significant_vars <- subset(female_aov_results_df, p_value_adj < 0.05)
new_row_names <- gsub("\\.1$", "", rownames(female_significant_vars))
rownames(female_significant_vars) <- new_row_names

print(female_significant_vars)

```


## Grouping by Group

### Normality by Group

```{r}
# Perform Shapiro-Wilk test for each numeric variable grouped by "Group" 
aovdata <- data[, c(3,6, 8:17)]
shapiro_test_group <- aovdata %>%
  group_by(Group) %>%
  summarize(across(where(function(x) length(unique(x)) > 1 && is.numeric(x)), ~ shapiro.test(.)$p.value, .names = "{.col}")) %>%
  ungroup()
shapiro_test_group<-as.data.frame(t(shapiro_test_group))

# Rename columns V1 and V2 to Females and Males
colnames(shapiro_test_group)[colnames(shapiro_test_group) == "V1"] <- "NTC"
colnames(shapiro_test_group)[colnames(shapiro_test_group) == "V2"] <- "WT"
colnames(shapiro_test_group)[colnames(shapiro_test_group) == "V3"] <- "L115R"
colnames(shapiro_test_group)[colnames(shapiro_test_group) == "V4"] <- "L116"

# Format the values in scientific notation to regular numeric format
shapiro_test_group$NTC <- format(as.numeric(shapiro_test_group$NTC), scientific = FALSE)
shapiro_test_group$WT <- format(as.numeric(shapiro_test_group$WT), scientific = FALSE)
shapiro_test_group$L115R <- format(as.numeric(shapiro_test_group$L115R), scientific = FALSE)
shapiro_test_group$L116 <- format(as.numeric(shapiro_test_group$L116), scientific = FALSE)


# Add a new column for the test result
shapiro_test_group$Significance <- ifelse(shapiro_test_group$NTC < 0.05|shapiro_test_group$WT< 0.05|shapiro_test_group$L115R <0.05|shapiro_test_group$L116<0.05, "Non-Normal", "Normal")
shapiro_test_group$Significance_NTC <- ifelse(shapiro_test_group$NTC < 0.05, "Non-Normal", "Normal")
shapiro_test_group$Significance_WT <- ifelse(shapiro_test_group$WT < 0.05, "Non-Normal", "Normal")
shapiro_test_group$Significance_L115R <- ifelse(shapiro_test_group$L115R < 0.05, "Non-Normal", "Normal")
shapiro_test_group$Significance_L116 <- ifelse(shapiro_test_group$L116 < 0.05, "Non-Normal", "Normal")

# Remove the "Gender" row
shapiro_tes_group <- shapiro_test_group[!(row.names(shapiro_test_group) %in% c("Group")),]
# Arrange the values by Females in ascending order
shapiro_test_group <- shapiro_test_group[order(shapiro_test_group$NTC), ]
# Add row names as a new column to the data frame
shapiro_test_group <- shapiro_test_group %>% rownames_to_column(var = "Variable")
# View the result
print(shapiro_test_group)

# Get variables that are not significant for both Males and Females
not_significant_group <- shapiro_test_group %>%
  filter(Significance == "Non-Normal") %>% 
  pull(Variable)
# Get variables that are not significant for both Males and Females
significant_group <- shapiro_test_group %>%
  filter(Significance == "Normal") %>% 
  pull(Variable)
```
## Mixed ANOVA

```{r}
# Assuming your original data frame is named 'data', select relevant columns
aovdata <- data[, c(3,6, 8:17)]

# Convert "Group" and "Gender" to factors
aovdata$Group <- as.factor(aovdata$Group)
aovdata$Gender <- as.factor(aovdata$Gender)

# Initialize an empty list to store ANOVA or Kruskal-Wallis results
aovlist <- list()


for (col in names(aovdata)[3:ncol(aovdata)]) {
      variable <- col
    formula <- as.formula(paste0("`", col, "`"," ~ Group"))
  
  # Check for normality assumption using Shapiro-Wilk test
  if (shapiro.test(aovdata[[col]])$p.value >= 0.05) {
    #Include Levene Test here
    res.aov <- aov(formula, data = aovdata)
    data_input <- data.frame("Group"=aovdata$Group, "Gender"=aovdata$Gender,aovdata[[col]])
    names(data_input)[3]<- variable
    data_input$Group <- as.factor(data_input$Group)
    posthoc <- data_input %>% rstatix::tukey_hsd(formula) # Tukey's post-hoc test for ANOVA
    posthoc<- posthoc %>% add_xy_position()
    test_type <- "ANOVA"
    Group <- aovdata$Group
    Group1 <- posthoc$group1
    Group2 <- posthoc$group2
    #pair_compared <- rownames(posthoc$`Group`)
  } else {
    res.aov <- kruskal.test(formula, data = aovdata)
    data_input <- data.frame("Group"=aovdata$Group, "Gender"=aovdata$Gender,aovdata[[col]])
    names(data_input)[3]<- variable
    data_input$Group <- as.factor(data_input$Group)
    posthoc <- data_input %>% dunn_test(formula)
    test_type <- "Kruskal-Wallis"
    Group <- aovdata$Group
    Group1 <- posthoc$group1
    Group2 <- posthoc$group2
    posthoc<- posthoc %>% add_xy_position()
    #pair_compared <- posthoc$comparison
  }
 aovlist[[col]] <- list(aov_results = broom::tidy(res.aov, conf.int = TRUE), Group = Group, posthoc_results = posthoc, Test_Type = test_type, Variable = variable)
}
```


```{r}
# Create empty data frame to store the results
aov_results_df <- data.frame()

# Loop through the list of test results
for (i in seq_along(aovlist)) {
  test_result <- aovlist[[i]]
   # Extract variable name
  variable <- test_result$Variable
  # Extract test type
  test_type <- test_result$Test_Type
  # Extract comparison and p-values
  if (grepl("ANOVA", test_type, ignore.case = TRUE)) {
    comparisons <- data.frame("Comparison"=paste0(test_result$posthoc_results$group1,"-",test_result$posthoc_results$group2),
                              "p_value_adj"=test_result$posthoc_results$p.adj,
                              "p.adj.signif"=test_result$posthoc_results$p.adj.signif,
                              "y.position"=test_result$posthoc_results$y.position,
                              "xmin"=test_result$posthoc_results$xmin,
                              "xmax" = test_result$posthoc_results$xmax)
    p_values <- test_result$aov_results$p.value[1]
    
    
  } else if (grepl("Kruskal-Wallis", test_type, ignore.case = TRUE)) {
    comparisons <- data.frame("Comparison"=paste0(test_result$posthoc_results$group1,"-",test_result$posthoc_results$group2),
                              "p_value_adj"=test_result$posthoc_results$p.adj,
                              "p.adj.signif"=test_result$posthoc_results$p.adj.signif,
                              "y.position"=test_result$posthoc_results$y.position,
                              "xmin"=test_result$posthoc_results$xmin,
                              "xmax" = test_result$posthoc_results$xmax)
    p_values <- test_result$aov_results$p.value
  }
  
  # Combine the results into a data frame
  aov_results_df <- rbind(aov_results_df, data.frame(
    Variable = variable,
    Test_Type = test_type,
    Comparison = comparisons,
    P_Value = p_values
  ))
}
#Add StarScale
#male_aov_results_df$StarScale <- sapply(male_aov_results_df$Comparison.p_value_adj, convert_to_star)
#Add gender column
aov_results_df$Gender <- "Male"


# Usamos strsplit para dividir la columna por el carácter "-" o " - "
split_data <- strsplit(aov_results_df$Comparison.Comparison, split = "-")


# Extraemos los elementos de la lista resultante y los asignamos a las nuevas columnas
aov_results_df$group1 <- sapply(split_data, function(x) x[[1]])
aov_results_df$group2 <- sapply(split_data, function(x) x[[2]])
#change names of columns
colnames(aov_results_df) <- gsub("Comparison\\.", "", colnames(aov_results_df))

aov_results_df<- aov_results_df %>% mutate_if(is.character, as.factor)
# Print the dataframe
print(aov_results_df)

# Filter variables with significant differences between groups based on the ANOVA results (alpha level = 0.05)
significant_vars <- subset(aov_results_df, p_value_adj < 0.05)
new_row_names <- gsub("\\.1$", "", rownames(significant_vars))
rownames(significant_vars) <- new_row_names

print(significant_vars)

```



## ANOVA Males vs Females
```{r}
# Assuming your original data frame is named 'data', select relevant columns
aovdata <- data[, c(3,6, 8:17)]

# Convert "Group" and "Gender" to factors
aovdata$Group <- as.factor(aovdata$Group)
aovdata$Gender <- as.factor(aovdata$Gender)

# Initialize an empty list to store ANOVA or Kruskal-Wallis results
aovlist <- list()
grupos <- c("NTC","WT","L115R","L116")

for(i in 1:length(grupos)){
mixed_data <- aovdata %>% filter(Group == grupos[i])
  #males_data <-mixed_data %>%  filter(Gender == "Male") %>%  select(names(mixed_data)[3:ncol(mixed_data)])
  #females_data <-mixed_data %>%  filter(Gender == "Female") %>%  select(names(mixed_data)[3:ncol(mixed_data)])
  group<- grupos[i]
   # Create a new list for each group
  group_result <- list()
for (col in names(mixed_data)[3:ncol(mixed_data)]) {
  formula <- as.formula(paste0("`", col, "`"," ~ Gender"))
  # Check for normality assumption using Shapiro-Wilk test
  if (shapiro.test(mixed_data[[col]])$p.value >= 0.05) {
    #Include Levene Test here
    #group<- grupos[i]
    variable <- col
    t_test_result <- mixed_data%>% t_test(formula) %>% add_significance()
    #test_type <- "T-Test"
    #t_test_result<- t_test_result %>% add_x_position(dodge = 0.8)
    t_test_result<- t_test_result %>% add_y_position()
    t_test_result <- t_test_result %>% add_x_position()
    
    t_test_result <- t_test_result %>% mutate(Test_type = "T-Test",
                                              Group_dif = grupos[i],
                                              xmin = case_when(
                                                Group_dif == "NTC"~0.8,
                                                Group_dif == "WT"~1.8,
                                                Group_dif == "L115R"~2.8,
                                                Group_dif == "L116"~3.8),
                                              xmax = case_when(
                                                Group_dif == "NTC"~1.2,
                                                Group_dif == "WT"~2.2,
                                                Group_dif == "L115R"~3.2,
                                                Group_dif == "L116"~4.2))
    #Modif x position
  } else {
    #group<- grupos[i]
    variable <- col
    t_test_result <- mixed_data%>% wilcox_test(formula) %>% add_significance()
    #test_type <- "Mann-Whitney"
    t_test_result <- t_test_result %>% add_y_position()
    t_test_result <- t_test_result %>% add_x_position()
    
    t_test_result <- t_test_result %>% mutate(Test_type = "Mann-Whitney",
                                              Group_dif = grupos[i],
                                              df = "df",
                                              xmin = case_when(
                                                Group_dif == "NTC"~0.8,
                                                Group_dif == "WT"~1.8,
                                                Group_dif == "L115R"~2.8,
                                                Group_dif == "L116"~3.8),
                                              xmax = case_when(
                                                Group_dif == "NTC"~1.2,
                                                Group_dif == "WT"~2.2,
                                                Group_dif == "L115R"~3.2,
                                                Group_dif == "L116"~4.2))
    #Modif x position
  }
  #group_result[[col]] <- list(results = t_test_result,Test_Type = test_type, Group= group)
  group_result[[col]] <- list(results = t_test_result)
  
}
  aovlist[[group]] <- group_result
}
```



```{r}
# Create empty data frame to store the results
aov_results_df_sex <- data.frame()
# Loop through the list of test results
for (i in seq_along(aovlist)) {
  group <- aovlist[[i]]
  
  for (j in seq_along(group)){
  variable <- group[[j]]$results
  # Append the row to the aov_results_df
  aov_results_df_sex <- rbind(aov_results_df_sex, variable)
}
}
# Print the dataframe
print(aov_results_df_sex)

# Filter variables with significant differences between groups based on the ANOVA results (alpha level = 0.05)
significant_vars_sex <- subset(aov_results_df_sex, p < 0.05)
new_row_names <- gsub("\\.1$", "", rownames(significant_vars_sex))
rownames(significant_vars_sex) <- new_row_names
print(significant_vars_sex)
```
### Combine by sex results

```{r}
#Combine both dataframes

both_aov_results_df <- rbind(male_aov_results_df, female_aov_results_df)
# Assuming 'aov_results_df' is the data frame you have
desired_order <- unique(aov_results_df$Variable)
# Convert the "Variable" column to a factor with the desired order
both_aov_results_df$Variable <- factor(both_aov_results_df$Variable, levels = desired_order)
##Attach statistical differences between gender
#Change name of columns
```


```{r}

names(aov_results_df_sex) <- c("Variable","group1","group2","n1","n2","P_Value","df","p_value_adj", "p.adj.signif", "y.position", "Comparison","xmin" ,"xmax","Test_Type", "Group_dif")
# Change "Comparison" value from c("Female","Male") to Male-Female
aov_results_df_sex$Comparison<- "Male-Female"
#Change positions of columnas

selected_cols_bysex <- aov_results_df_sex %>% 
  select(Variable, Test_Type, Comparison, p_value_adj, p.adj.signif, y.position ,xmin, xmax,P_Value, Group_dif, group1, group2)

both_aov_results_df <- bind_rows(both_aov_results_df,selected_cols_bysex)


both_aov_results_df
# Filter variables with significant differences between groups based on the ANOVA results (alpha level = 0.05)
both_significant_vars <- subset(both_aov_results_df, p_value_adj < 0.05)
new_row_names <- gsub("\\.1$", "", rownames(both_significant_vars))
rownames(both_significant_vars) <- new_row_names

both_significant_vars

```



# Plots by sex

```{r}
# Combine male_varsnames and female_varsnames into varsplot
varsplot <- colnames(aovdata[,3:12])
#Compute comparison groups
my_comparisons <- list(c("NTC","WT"),c("NTC","L115R"),c("NTC","L116"),c("WT","L115R"),c("WT","L116"),c("L115R","L116"))
# Create a directory to save the plots (if it doesn't exist)
if (!file.exists("plots")) {
  dir.create("plots")
}

# Assuming 'aov_results_df' is the data frame you have
desired_order <- unique(aov_results_df$Variable)

# Convert the "Variable" column to a factor with the desired order
both_aov_results_df$Variable <- factor(both_aov_results_df$Variable, levels = desired_order)
both_dataframe_list <- both_aov_results_df %>%  group_split(Variable)

# Boxplot

suppressWarnings(
for (i in 1:length(varsplot)){
#data_males <- as.data.frame(males_dataframe_list[[i]])
#data_females <- as.data.frame(females_dataframe_list[[i]])
data_both <- as.data.frame(both_dataframe_list[[i]])  
#data_both_significant <- as.data.frame(both_significant_list[[i]])
if(all(data_both$p.adj.signif=="ns")){
p<-ggplot(data=aovdata, aes(x = Group, 
                            y = .data[[varsplot[i]]], 
                            fill = Gender)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.2), alpha = 0.9, size = 2) +  # Separate jittered points by "Gender"
  labs(x = "Group", 
       y = varsplot[i]) +
  ggtitle(paste0("Boxplot of ", varsplot[i])) +
  theme_minimal()+
  coord_cartesian(ylim = c(0, NA))

}else{
  p<-ggplot(data=aovdata, aes(x = Group, 
                            y = .data[[varsplot[i]]], 
                            fill = Gender)) +
  geom_boxplot(outlier.shape = NA) +
  geom_jitter(position = position_jitterdodge(jitter.width = 0.2), alpha = 0.9, size = 2) +  # Separate jittered points by "Gender"
  labs(x = "Group", 
       y = varsplot[i]) +
  ggtitle(paste0("Boxplot of ", varsplot[i])) +
  theme_minimal()+
  coord_cartesian(ylim = c(0, NA))+
    stat_pvalue_manual(data_both, 
                       label = "p.adj.signif", 
                       hide.ns=TRUE, 
                       y.position = "y.position", 
                       step.increase = 0.1,
                       xmin = "xmin", 
                       xmax = "xmax", 
                       dodge=0.8)
}

print(p)
# File path to save the plot
  file_name_cleaned <- clean_variable_name(varsplot[i])
  file_path <- paste0("plots/", file_name_cleaned, "_boxplot.pdf")
  
  # Save the plot in vector format (PDF)
  ggsave(filename = file_path, plot = p, device = "pdf", width = 8, height = 6, units = "in")
})
```



# Plots mixed

```{r}
# Combine male_varsnames and female_varsnames into varsplot
varsplot <- colnames(aovdata[,3:12])
#Compute comparison groups
my_comparisons <- list(c("NTC","WT"),c("NTC","L115R"),c("NTC","L116"),c("WT","L115R"),c("WT","L116"),c("L115R","L116"))

# Create a directory to save the plots (if it doesn't exist)
if (!file.exists("plots")) {
  dir.create("plots")
}

# Assuming 'aov_results_df' is the data frame you have
desired_order <- unique(aov_results_df$Variable)

# Convert the "Variable" column to a factor with the desired order
aov_results_df$Variable <- factor(aov_results_df$Variable, levels = desired_order)

# Group the data frame by "Variable" and store the groups in a list
grouped_list <- aov_results_df %>%  group_split(Variable)
```


```{r}
# Boxplot
for (i in 1:length(varsplot)){
  data_mixed <- as.data.frame(grouped_list[[i]])
    if(all(data_mixed$p.adj.signif=="ns")){
        p<-ggplot(data=aovdata, aes(x = Group, 
                            y = .data[[varsplot[i]]],
                            fill = Group)) +
            geom_boxplot(outlier.shape = NA) +
            geom_jitter(aes(color= Gender), position = position_jitterdodge(jitter.width = 0.5),alpha = 0.9, size = 2) +  # Separate jittered points by "Gender"
            labs(x = "Group", 
                 y = varsplot[i]) +
            ggtitle(paste0("Boxplot of ", varsplot[i])) +
            scale_fill_manual(values = c(NTC = "#808080",
                                         WT = "#D4D3D3",
                                         L115R = "#E42320",
                                         L116 = "#2A4B9A")) +
            theme_minimal()+
            coord_cartesian(ylim = c(0, NA))
    }else{
        p<-ggplot(data=aovdata, aes(x = Group, 
                            y = .data[[varsplot[i]]],
                            fill = Group)) +
            geom_boxplot(outlier.shape = NA) +
            geom_jitter(aes(color= Gender), position = position_jitterdodge(jitter.width = 0.5),alpha = 0.9, size = 2) +  # Separate jittered points by "Gender"
            labs(x = "Group", 
                 y = varsplot[i]) +
            ggtitle(paste0("Boxplot of ", varsplot[i])) +
            scale_fill_manual(values = c(NTC = "#808080",
                                         WT = "#D4D3D3",
                                         L115R = "#E42320",
                                         L116 = "#2A4B9A")) +
            theme_minimal()+
            coord_cartesian(ylim = c(0, NA))+
            stat_pvalue_manual( data = data_mixed, 
                                mapping = aes(xmin = xmin, xmax = xmax, y.position = y.position, label = p.adj.signif),
                               label = "p.adj.signif",
                               hide.ns = TRUE,
                               step.increase = 0.1,
                               inherit.aes = FALSE) +
          labs(fill = "Group")  # This is important to show the correct legend
  
}
print(p)
# File path to save the plot
  file_name_cleaned <- clean_variable_name(varsplot[i])
  file_path <- paste0("plots/", file_name_cleaned, "_Mixed_boxplot.pdf")
  
  # Save the plot in vector format (PDF)
  ggsave(filename = file_path, plot = p, device = "pdf", width = 8, height = 6, units = "in")
}
```
```{r}
    data_mixed <- as.data.frame(grouped_list[[1]])
```
# Correlation between number of falls and time to fall with weight

```{r}
# Create scatter plots with tendency lines


scatter_plot <- data %>%
  group_by(Gender, Group) %>%
  mutate(R2 = summary(lm(Mean_time ~ Weight_in_g))$r.squared) %>%
  ggplot(aes(x = Weight_in_g, y = Mean_time, color = Group)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line
  geom_text(
    aes(label = paste("R^2 =", round(R2, 2))),
    x = Inf, y = -Inf, hjust = 1, vjust = 0, show.legend = FALSE
  ) +
  facet_grid(Gender ~ .) +  # Separate plots for each sex
  labs(x = "Weight (g)", y = "Mean Time", color = "Group") +
  theme_minimal()+
  ggtitle("Mean_time_by_weight and sex")

# Display the scatter plot
print(scatter_plot)
  # Save the plot in vector format (PDF)
  ggsave(filename = "Mean_time_by_weight and sex.pdf", plot = scatter_plot, device = "pdf", width = 8, height = 6, units = "in")
```
```{r}
# Create scatter plots with tendency lines


scatter_plot <- data %>%
  group_by(Gender, Group) %>%
  mutate(R2 = summary(lm(Mean_falls ~ Weight_in_g))$r.squared) %>%
  ggplot(aes(x = Weight_in_g, y = Mean_falls, color = Group)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line
  geom_text(
    aes(label = paste("R^2 =", round(R2, 2))),
    x = Inf, y = -Inf, hjust = 1, vjust = 0, show.legend = FALSE
  ) +
  facet_grid(Gender ~ .) +  # Separate plots for each sex
  labs(x = "Weight (g)", y = "Mean Falls", color = "Group") +
  theme_minimal()+
  ggtitle("Mean_falls_by_weight and sex")

# Display the scatter plot
print(scatter_plot)
  # Save the plot in vector format (PDF)
  ggsave(filename = "Mean_falls_by_weight and sex.pdf", plot = scatter_plot, device = "pdf", width = 8, height = 6, units = "in")
```



```{r}
# Create scatter plots with tendency lines


scatter_plot <- data %>%
  group_by(Group) %>%
  mutate(R2 = summary(lm(Mean_time ~ Weight_in_g))$r.squared) %>%
  ggplot(aes(x = Weight_in_g, y = Mean_time, color = Group)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line
  geom_text(
    aes(label = paste("R^2 =", round(R2, 2))),
    x = Inf, y = -Inf, hjust = 1, vjust = 0, show.legend = FALSE
  ) +
  labs(x = "Weight (g)", y = "Mean Time", color = "Group") +
  theme_minimal()+
  ggtitle("Mean_time_by_weight")

# Display the scatter plot
print(scatter_plot)
  # Save the plot in vector format (PDF)
  ggsave(filename = "Mean_time_by_weight.pdf", plot = scatter_plot, device = "pdf", width = 8, height = 6, units = "in")
```
```{r}
# Create scatter plots with tendency lines
scatter_plot <- data %>%
  group_by(Group) %>%
  mutate(R2 = summary(lm(Mean_falls ~ Weight_in_g))$r.squared) %>%
  ggplot(aes(x = Weight_in_g, y = Mean_falls, color = Group)) +
  geom_point() +
  geom_smooth(method = "lm", se = FALSE) +  # Add linear regression line
  geom_text(
    aes(label = paste("R^2 =", round(R2, 2))),
    x = Inf, y = -Inf, hjust = 1, vjust = 0, show.legend = FALSE
  ) +
  labs(x = "Weight (g)", y = "Mean Falls", color = "Group") +
  theme_minimal()+
  ggtitle("Mean_falls_by_weight")

# Display the scatter plot
print(scatter_plot)
  # Save the plot in vector format (PDF)
  ggsave(filename = "Mean_falls_by_weight.pdf", plot = scatter_plot, device = "pdf", width = 8, height = 6, units = "in")
```


# Dataframes to save

```{r}
#Group by Gender
levene_results_gender
shapiro_test_gender

male_aov_results_df
male_significant_vars
female_aov_results_df
female_significant_vars
#Mixed ANOVA
shapiro_test_group
aov_results_df
significant_vars
#Males vs Females
aov_results_df_sex
significant_vars_sex

```
# Tables
```{r}
# Load required libraries
library(kableExtra)
dataframes <- c("levene_results_gender","shapiro_test_gender","male_aov_results_df","male_significant_vars","female_aov_results_df","female_significant_vars","shapiro_test_group","aov_results_df","significant_vars","aov_results_df_sex","significant_vars_sex")

for (i in 1:length(dataframes)) {
  df <- get(dataframes[i])
  cols <- colnames(df)  # Get the column names of the dataframe
  p <- knitr::kable(df, "html") %>% 
    kable_paper("striped", full_width = F, font_size = 14) 
  #%>%     add_header_above(header = list(cols))  # Use a list to specify the header row
  print(dataframes[i])
  print(p)
}

```
# Export resulst as CSV
```{r}

# Carga la librería openxlsx
library(openxlsx)

# Crea un nuevo libro de Excel
wb <- createWorkbook()

# Agrega cada dataframe como una hoja en el libro
for (df_name in dataframes) {
  df <- get(df_name)  # Get the dataframe by its name
  addWorksheet(wb, sheetName = df_name)
  writeData(wb, sheet = df_name, x = df)
}

# Guarda el libro de Excel
saveWorkbook(wb, file = "dataframes.xlsx", overwrite = TRUE)
```




