---
title: "Survival"
date: "2023-09-04"
author: Santiago López Begines
github: https://github.com/SLopezBegines/R_code
output:
  html_document:
    highlight: kate
    fig_width: 8
    fig_height: 7
    fig_caption: no
    lightbox: yes
    gallery: yes
    use_bookdown: yes
    code_folding: show
    toc: yes
    toc_depth: 5
    toc_float: 
      collapsed: no
      smooth_scroll: yes
    df_print: kable
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

```{r}
#Libraries
library(readxl)
library(tidyverse)
library(psych)
library(DescTools)
library(car)
library(stats)
library(broom)
library(dunn.test)
library(stats)
library(ggpubr)
library(RcmdrMisc)
library(spatstat.utils)
```


# 1. Load data

```{r}
data<- read_excel("Survival_all.xlsx")
```

# 2. Cleaning data
Clean the data. Data includes all values with all types of deaths. Age is calculated in months and days. 
```{r}
#Round age in months. 
data$Age_m <- round(data$Age_m, 0)
data$Group <- factor(data$Group, levels=c("NTC","WT","L115","L116"))

#Remove animals with more than 30months and less than 1month.
data <- data[data$Age_m>=1 & data$Age_m<=30,]
```

Statistical summary of data
```{r}
mean_sem_data <- data %>% 
  filter(Type_of_dead == "Natural Death") %>% 
  group_by(Gender, Group) %>% 
  summarise(mean_m=mean(Age_m,na.rm = TRUE),
            SD=sd(Age_m,na.rm = TRUE),
            SEM=SD/sqrt(n()),
            N=n(),
            mean_d=mean(Age_d,na.rm = TRUE),
            SD_d=sd(Age_d,na.rm = TRUE),
            SEM_d=SD/sqrt(n()),
            MinAge_m=min(Age_m),
            MaxAge_m=max(Age_m),
            MinAge_d=min(Age_d),
            MaxAge_d=max(Age_d))
mean_sem_data
```


# 3. Mixed Data
Mixing males and females. Taking only account "Natural Deaths"
```{r}
mn<-data %>% 
  filter(Type_of_dead=="Natural Death")

p<- ggplot(mn, aes(x = Age_m, color = Group)) +
  stat_ecdf(geom = "step", size=0.5) +
  scale_color_manual(values = c(NTC = "#808080",
                               WT = "#000000",
                               L115 = "#E42320",
                               L116 = "#2A4B9A")) +
  labs(title = "Cumulative Distribution Comparison",
       x = "Age in Months",
       y = "Cumulative Probability") +
  theme_minimal()+
  theme(axis.line.x.bottom = element_line(color = "black", size = 0.5),
                  axis.line.y.left = element_line(color = "black", size = 0.5))
print(p)
# Create a directory to save the plots (if it doesn't exist)
if (!file.exists("plots")) {
  dir.create("plots")
}
# File path to save the plot
  file_path <- paste0("plots/Survival_Mixed_boxplot.pdf")
  
  # Save the plot in vector format (PDF)
  ggsave(filename = file_path, plot = p, device = "pdf", width = 8, height = 6, units = "in")

```

```{r}
# Load the ks package if not already loaded
library(ks)

mn<-data %>% 
  filter(Type_of_dead=="Natural Death")
# Create an empty matrix to store the p-values
p_values_matrix <- matrix(NA, nrow = 4, ncol = 4)

# Get unique group names
unique_groups <- unique(mn$Group)

# Loop through all possible pairwise combinations of groups
for (i in 1:(length(unique_groups) - 1)) {
  for (j in (i + 1):length(unique_groups)) {
    group_1 <- unique_groups[i]
    group_2 <- unique_groups[j]
    # Subset the data for the two groups being compared
    #data_group1 <- MN_data[MN_data$Group == group1, "cum"]
    #data_group2 <- MN_data[MN_data$Group == group2, "cum"]
    data_group_1 <- mn %>% filter(Group==group_1) %>% pull(Age_m)
    data_group_2 <- mn %>% filter(Group==group_2) %>% pull(Age_m)
    # Perform the KS test
    ks_result <- ks.test(data_group_1, data_group_2)
    
    # Store the p-value in the matrix
    p_values_matrix[i, j] <- ks_result$p.value
    p_values_matrix[j, i] <- ks_result$p.value
  }
}

# Add row and column names to the p-values matrix
rownames(p_values_matrix) <- unique_groups
colnames(p_values_matrix) <- unique_groups
p_val <- as.data.frame(p_values_matrix)
# View the p-values matrix (optional)
print(p_val)
```

Comparison between WT and L115R
```{r}
mn<-data %>% 
  filter(Type_of_dead=="Natural Death"& Group %in% c("WT","L115"))

p<- ggplot(mn, aes(x = Age_m, color = Group)) +
  stat_ecdf(geom = "step", size=0.5) +
  scale_color_manual(values = c(NTC = "#808080",
                               WT = "#000000",
                               L115 = "#E42320",
                               L116 = "#2A4B9A")) +
  labs(title = "Cumulative Distribution Comparison",
       x = "Age in Months",
       y = "Cumulative Probability") +
  theme_minimal()+
  theme(axis.line.x.bottom = element_line(color = "black", size = 0.5),
                  axis.line.y.left = element_line(color = "black", size = 0.5))
print(p)
# File path to save the plot
  file_path <- paste0("plots/Survival_Mixed_boxplot_WT_L115.pdf")
  
  # Save the plot in vector format (PDF)
  ggsave(filename = file_path, plot = p, device = "pdf", width = 8, height = 6, units = "in")
```
### Kaplan-Meier Plots
All types of deaths are taken into account. Status variable is created. Status is set as 0 when age have been censored, and 1 when death occured before to end the experiment (Natural death are marked as 1). Censored are the samples that have been sacrified in any way before their natural death. 
```{r}
library(ggsurvfit)
library(gtsummary)
library(tidycmprsk)


survival <- survival::Surv(data$Age_m, data$Status)
s1<- survival::survfit(survival~Group, data)

p<- survfit2(Surv(Age_m, Status) ~ Group, data ) %>% 
  ggsurvfit() +
  labs(
    x = "Months",
    y = "Overall survival probability"
  )+
  theme_minimal()+
  theme(axis.line.x.bottom = element_line(color = "black", size = 0.5),
                  axis.line.y.left = element_line(color = "black", size = 0.5))+
    scale_color_manual(values = c(NTC = "#808080",
                               WT = "#000000",
                               L115 = "#E42320",
                               L116 = "#2A4B9A"))
print(p)
# File path to save the plot
  file_path <- paste0("plots/Survival_KM_Mixed_boxplot.pdf")
  
  # Save the plot in vector format (PDF)
  ggsave(filename = file_path, plot = p, device = "pdf", width = 8, height = 6, units = "in")
```
#### Long-Rank test
p-value of Long-Rank test is p=0.5, so null hypothesis is not rejected. There are not differences between genotypes.

```{r}
long_rank_mixed <- survival::survdiff(Surv(Age_m, Status) ~ Group, data)
long_rank_mixed
```
```{r}
survival::coxph(Surv(Age_m,Status) ~ Gender, data) %>% 
  tbl_regression(exp=TRUE)
```


# 4. Females
Taking only account Females and "Natural Deaths"
```{r}
mn<-data %>% 
  filter(Type_of_dead=="Natural Death" & Gender =="Female")

p<- ggplot(mn, aes(x = Age_m, color = Group)) +
  stat_ecdf(geom = "step", size=0.5) +
  scale_color_manual(values = c(NTC = "#808080",
                               WT = "#000000",
                               L115 = "#E42320",
                               L116 = "#2A4B9A")) +
  labs(title = "Cumulative Distribution Comparison",
       x = "Age in Months",
       y = "Cumulative Probability") +
  theme_minimal()+theme(axis.line.x.bottom = element_line(color = "black", size = 0.5),
                  axis.line.y.left = element_line(color = "black", size = 0.5))
print(p)
# Create a directory to save the plots (if it doesn't exist)
if (!file.exists("plots")) {
  dir.create("plots")
}
# File path to save the plot
  file_path <- paste0("plots/Survival_Females_boxplot.pdf")
  
  # Save the plot in vector format (PDF)
  ggsave(filename = file_path, plot = p, device = "pdf", width = 8, height = 6, units = "in")

```

```{r}
# Load the ks package if not already loaded
library(ks)

mn<-data %>% 
  filter(Type_of_dead=="Natural Death"& Gender =="Female")
# Create an empty matrix to store the p-values
p_values_matrix <- matrix(NA, nrow = 4, ncol = 4)

# Get unique group names
unique_groups <- unique(mn$Group)

# Loop through all possible pairwise combinations of groups
for (i in 1:(length(unique_groups) - 1)) {
  for (j in (i + 1):length(unique_groups)) {
    group_1 <- unique_groups[i]
    group_2 <- unique_groups[j]
    # Subset the data for the two groups being compared
    #data_group1 <- MN_data[MN_data$Group == group1, "cum"]
    #data_group2 <- MN_data[MN_data$Group == group2, "cum"]
    data_group_1 <- mn %>% filter(Group==group_1) %>% pull(Age_m)
    data_group_2 <- mn %>% filter(Group==group_2) %>% pull(Age_m)
    # Perform the KS test
    ks_result <- ks.test(data_group_1, data_group_2)
    
    # Store the p-value in the matrix
    p_values_matrix[i, j] <- ks_result$p.value
    p_values_matrix[j, i] <- ks_result$p.value
  }
}

# Add row and column names to the p-values matrix
rownames(p_values_matrix) <- unique_groups
colnames(p_values_matrix) <- unique_groups
p_val_female<- as.data.frame(p_values_matrix)
# View the p-values matrix (optional)
print(p_val_female)
```

### Kaplan-Meier Plots
All types of deaths are taken into account. Status variable is created. Status is set as 0 when age have been censored, and 1 when death occured before to end the experiment (Natural death are marked as 1). Censored are the samples that have been sacrified in any way before their natural death. 
```{r}
library(ggsurvfit)
library(gtsummary)
library(tidycmprsk)

km_data_female <- data %>% 
  filter(Gender =="Female")

survival <- survival::Surv(km_data_female$Age_m, km_data_female$Status)
s1<- survival::survfit(survival~Group, km_data_female)

p<- survfit2(Surv(Age_m, Status) ~ Group, km_data_female) %>% 
  ggsurvfit() +
  labs(
    x = "Months",
    y = "Overall survival probability"
  )+
  theme_minimal()+
  theme(axis.line.x.bottom = element_line(color = "black", size = 0.5),
                  axis.line.y.left = element_line(color = "black", size = 0.5))+
    scale_color_manual(values = c(NTC = "#808080",
                               WT = "#000000",
                               L115 = "#E42320",
                               L116 = "#2A4B9A"))
print(p)
# File path to save the plot
  file_path <- paste0("plots/Survival_KM_Females_boxplot.pdf")
  
  # Save the plot in vector format (PDF)
  ggsave(filename = file_path, plot = p, device = "pdf", width = 8, height = 6, units = "in")
```
#### Long-Rank test
p-value of Long-Rank test is p=0.1, so null hypothesis is not rejected. There are not differences between genotypes.


```{r}
survival::survdiff(Surv(Age_m, Status) ~ Group, km_data_female)
```
```{r}
survival::coxph(Surv(Age_m,Status) ~ Group, km_data_female) %>% 
  tbl_regression(exp=TRUE)
```





# 5. Males
Taking only account Males and "Natural Deaths"

```{r}
mn<-data %>% 
  filter(Type_of_dead=="Natural Death" & Gender =="Male")

p<- ggplot(mn, aes(x = Age_m, color = Group)) +
  stat_ecdf(geom = "step", size=0.5) +
  scale_color_manual(values = c(NTC = "#808080",
                               WT = "#000000",
                               L115 = "#E42320",
                               L116 = "#2A4B9A")) +
  labs(title = "Cumulative Distribution Comparison",
       x = "Age in Months",
       y = "Cumulative Probability") +
  theme_minimal()+
  theme(axis.line.x.bottom = element_line(color = "black", size = 0.5),
                  axis.line.y.left = element_line(color = "black", size = 0.5))
print(p)
# Create a directory to save the plots (if it doesn't exist)
if (!file.exists("plots")) {
  dir.create("plots")
}
# File path to save the plot
  file_path <- paste0("plots/Survival_Males_boxplot.pdf")
  
  # Save the plot in vector format (PDF)
  ggsave(filename = file_path, plot = p, device = "pdf", width = 8, height = 6, units = "in")

```


```{r}
# Load the ks package if not already loaded
library(ks)

mn<-data %>% 
  filter(Type_of_dead=="Natural Death"& Gender =="Male")
# Create an empty matrix to store the p-values
p_values_matrix <- matrix(NA, nrow = 4, ncol = 4)

# Get unique group names
unique_groups <- unique(mn$Group)

# Loop through all possible pairwise combinations of groups
for (i in 1:(length(unique_groups) - 1)) {
  for (j in (i + 1):length(unique_groups)) {
    group_1 <- unique_groups[i]
    group_2 <- unique_groups[j]
    # Subset the data for the two groups being compared
    #data_group1 <- MN_data[MN_data$Group == group1, "cum"]
    #data_group2 <- MN_data[MN_data$Group == group2, "cum"]
    data_group_1 <- mn %>% filter(Group==group_1) %>% pull(Age_m)
    data_group_2 <- mn %>% filter(Group==group_2) %>% pull(Age_m)
    # Perform the KS test
    ks_result <- ks.test(data_group_1, data_group_2)
    
    # Store the p-value in the matrix
    p_values_matrix[i, j] <- ks_result$p.value
    p_values_matrix[j, i] <- ks_result$p.value
  }
}

# Add row and column names to the p-values matrix
rownames(p_values_matrix) <- unique_groups
colnames(p_values_matrix) <- unique_groups
p_val_male <- as.data.frame(p_values_matrix)
# View the p-values matrix (optional)
print(p_val_male)
```

```{r}
mn<-data %>% 
  filter(Type_of_dead=="Natural Death" & Gender =="Male" & Group %in% c("WT","L115","L116"))

p<- ggplot(mn, aes(x = Age_m, color = Group)) +
  stat_ecdf(geom = "step") +
  scale_color_manual(values = c(NTC = "#808080",
                               WT = "#000000",
                               L115 = "#E42320",
                               L116 = "#2A4B9A")) +
  labs(title = "Cumulative Distribution Comparison",
       x = "Age in Months",
       y = "Cumulative Probability") +
  theme_minimal()
print(p)
# Create a directory to save the plots (if it doesn't exist)
if (!file.exists("plots")) {
  dir.create("plots")
}
# File path to save the plot
  file_path <- paste0("plots/Survival?_Males_boxplot_WT_L115.pdf")
  
  # Save the plot in vector format (PDF)
  ggsave(filename = file_path, plot = p, device = "pdf", width = 8, height = 6, units = "in")

```


### Kaplan-Meier Plots
All types of deaths are taken into account. Status variable is created. Status is set as 0 when age have been censored, and 1 when death occured before to end the experiment (Natural death are marked as 1). Censored are the samples that have been sacrified in any way before their natural death. 
```{r}
library(ggsurvfit)
library(gtsummary)
library(tidycmprsk)

km_data_male <- data %>% 
  filter(Gender =="Male")

survival <- survival::Surv(km_data_male$Age_m, km_data_male$Status)
s1<- survival::survfit(survival~Group, km_data_male)

p<- survfit2(Surv(Age_m, Status) ~ Group, km_data_male) %>% 
  ggsurvfit() +
  labs(
    x = "Months",
    y = "Overall survival probability"
  )+
  theme_minimal()+
  theme(axis.line.x.bottom = element_line(color = "black", size = 0.5),
                  axis.line.y.left = element_line(color = "black", size = 0.5))+
    scale_color_manual(values = c(NTC = "#808080",
                               WT = "#000000",
                               L115 = "#E42320",
                               L116 = "#2A4B9A"))
print(p)
# File path to save the plot
  file_path <- paste0("plots/Survival_KM_Males_boxplot.pdf")
  
  # Save the plot in vector format (PDF)
  ggsave(filename = file_path, plot = p, device = "pdf", width = 8, height = 6, units = "in")
```
#### Long-Rank test
p-value of Long-Rank test is less p=0.05, so null hypothesis is rejected. There are differences between genotypes.

```{r}
survival::survdiff(Surv(Age_m, Status) ~ Group, km_data_male)
```
To check where are the differences between groups, post-hoc test is used. Differences only arise when "none" correction method is applied for correction of adjusted p-values.So, the difference exist only between L115 and the rest of the group, but this difference is no highly different statistically.

```{r}
survminer::pairwise_survdiff(Surv(Age_m, Status)~Group, km_data_male, p.adjust.method="none")
```

```{r}
survival::coxph(Surv(Age_m,Status) ~ Group, km_data_male) %>% 
  tbl_regression(exp=TRUE)
```
# 6. Tables
```{r}
# Load required libraries
library(kableExtra)
dataframes <- c("data","p_val","p_val_female","p_val_male", "mean_sem_data")

for (i in 1:length(dataframes)) {
  df <- get(dataframes[i])
  cols <- colnames(df)  # Get the column names of the dataframe
  p <- knitr::kable(df, "html") %>% 
    kableExtra::kable_paper("striped", full_width = F, font_size = 14)
  print(p)
}

```


# Export resulst as CSV
```{r}

# Carga la librería openxlsx
library(openxlsx)

# Crea un nuevo libro de Excel
wb <- createWorkbook()

# Agrega cada dataframe como una hoja en el libro
for (df_name in dataframes) {
  df <- get(df_name)  # Get the dataframe by its name
  addWorksheet(wb, sheetName = df_name)
  writeData(wb, sheet = df_name, x = df)
}

# Guarda el libro de Excel
saveWorkbook(wb, file = "dataframes.xlsx", overwrite = TRUE)
```
# 7. Session Info

```{r}
sessionInfo()
```




## Not included


```{r eval = FALSE}
MN_data <- data %>% 
  filter(Type_of_dead=="MN") %>% 
  group_by(Group, Age_m) %>% 
  summarise(N = n())
MN_data
```


```{r eval = FALSE}
MN_data$prop <- MN_data %>% 
  group_by(Group) %>% 
  summarise(prop=N/sum(N),
            cumprop=cumsum(prop))

MN_data$cum <- MN_data$prop$cumprop
MN_data$prop <- MN_data$prop$prop

```

```{r eval = FALSE}
ggplot(MN_data) +
  aes(x = Age_m, y = cum, fill = Group) +
  geom_col(position="dodge") +
  geom_line()+
  scale_fill_hue(direction = 1) +
    scale_fill_manual(values = c(NTC = "#808080",
                               WT = "#D4D3D3",
                               L115 = "#E42320",
                               L116 = "#2A4B9A")) +
  theme_minimal() 
```
